---
title: "Improvement in Mapping Rate in DNTx compared against Gencode v29"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook:
    theme: united
    toc: true
    toc_float: true
---

# Load Data
```{r}
library(pool)
library(tidyverse)
library(RSQLite)

gene_pool_2019 <- dbPool(drv = SQLite(), dbname = "/Volumes/ARC168/PROJECTS/mcgaughey/eyeIntegration_DNTx/EiaD_human_expression_2019_09.sqlite", idleTimeout = 3600000)
core_tight_2019 <- gene_pool_2019 %>% tbl('metadata') %>% as_tibble()
core_tight_2019$sample_accession<-gsub('E-MTAB-','E.MTAB.',core_tight_2019$sample_accession)
core_tight_2019$Sub_Tissue <- gsub('_',' - ',core_tight_2019$Sub_Tissue)
core_tight_2019 <- core_tight_2019 %>% select(-mapping_rate) %>% filter(!Tissue %in% c('Choroid Plexus','EyeLid'), !Sub_Tissue == 'Cornea - Stroma', !grepl("MGS", Sub_Tissue))


# fix tissue <-> color
meta <- 'core_tight_2019'
tissue_val <- setNames(c(pals::glasbey(n = 32), pals::kelly(n = get(meta) %>% pull(Tissue) %>% unique() %>% length() - 32)) %>% colorspace::lighten(0.3), get(meta) %>% pull(Tissue) %>% unique() %>% sort())
tissue_col <- scale_colour_manual(values = tissue_val)
tissue_fill <- scale_fill_manual(values = tissue_val)

# mapping rates
gencode <- read_delim('~/git/eyeIntegration_DNTx/data/mapping_rate_EiaD_2019_05.txt', 
                      col_names = c('sample_accession', 'mapping_rate'), 
                      delim = ' ') %>% 
  mutate(sample_accession = gsub('-', '.', sample_accession),
         mapping_rate = gsub('%','', mapping_rate) %>% as.numeric()) 
DNTx <- read_delim('~/git/eyeIntegration_DNTx/data/mapping_rates_EiaD_2019_09_DNTx_txome.txt', 
                   col_names = c('sample_accession', 'mapping_rate'), 
                   delim = ' ') %>% 
  mutate(sample_accession = gsub('-', '.', sample_accession),
         mapping_rate = gsub('%','', mapping_rate) %>% as.numeric()) 
```

# Mapping
Compare mapping rates against Gencode v29

## Distributions
Facetted by tissue, showing each Sub Tissue

We see a consistent improvement in mapping rate with DNTx
```{r, fig.width=20, fig.height=20}
bind_rows(core_tight_2019 %>% left_join(., gencode, by = 'sample_accession') %>% 
            mutate(fasta = 'Gencode v29'), 
          core_tight_2019 %>% left_join(., DNTx, by = 'sample_accession') %>% 
            mutate(fasta = 'DNTx 2019 09')) %>% 
  ggplot(aes(x=Sub_Tissue, y = mapping_rate, colour = fasta)) + 
  geom_violin(draw_quantiles = c(0.5)) + facet_wrap(~Tissue, scales = 'free_x') + 
  ggsci::scale_color_aaas() +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) 

```

## Delta Mean
Now let's simplify the visualization by cmoparing the DELTA mean (or median) between Sub_Tissues
```{r, fig.width=10, fig.height=10}
bind_rows(core_tight_2019 %>% left_join(., gencode, by = 'sample_accession') %>% 
            mutate(fasta = 'Gencode v29'), 
          core_tight_2019 %>% left_join(., DNTx, by = 'sample_accession') %>% 
            mutate(fasta = 'DNTx 2019 09')) %>% 
  group_by(Sub_Tissue, fasta) %>% 
  summarise(mean = mean(mapping_rate)) %>% 
  spread(fasta, mean) %>% 
  mutate(Delta = `DNTx 2019 09` - `Gencode v29`) %>% 
  left_join(., core_tight_2019 %>% select(Tissue, Sub_Tissue) %>% unique()) %>% 
  ggplot(aes(x=Sub_Tissue, y = Delta)) + geom_bar(stat = 'identity') +
  facet_wrap(~Tissue, scales = 'free_x') +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) 
```

## Delta Mean, no facetting, ordered by delta mean
Now let's simplify the visualization by cmoparing the DELTA mean (or median) between Sub_Tissues
```{r, fig.width=3, fig.height=4}
bind_rows(core_tight_2019 %>% left_join(., gencode, by = 'sample_accession') %>% 
            mutate(fasta = 'Gencode v29'), 
          core_tight_2019 %>% left_join(., DNTx, by = 'sample_accession') %>% 
            mutate(fasta = 'DNTx 2019 09')) %>% 
  group_by(Sub_Tissue, fasta) %>% 
  summarise(mean = mean(mapping_rate)) %>% 
  spread(fasta, mean) %>% 
  mutate(Delta = `DNTx 2019 09` - `Gencode v29`) %>% 
  left_join(., core_tight_2019 %>% select(Tissue, Sub_Tissue) %>% unique()) %>% 
  mutate(Group = case_when(Tissue %in% c('Cornea', 'ESC', 'EyeLid', 'Lens', 'Retina', 'Retinal Endothelium', 'RPE', 'Choroid Plexus') ~ 'Ocular',
                           TRUE ~ 'Body')) %>% 
  ggplot(aes(x=reorder(Sub_Tissue, Delta), y = Delta, fill = Group)) + geom_bar(stat = 'identity') +
  theme_minimal() +
  coord_flip() +
  ggsci::scale_fill_aaas() +
  xlab('EiaD Tissues') + ylab("Improvement in Mean Mapping\nPercentage Between Gencode V29 and DNTx V01")
```


```{r}
devtools::session_info()
```